The Economics of Shotgun Marriage

### 2020-10-02 slides-oct-20/slides.tex
Current stage

\setbeamertemplate{footline}[frame number]

@@

\begin{frame}
\frametitle{``America's strongest anti-poverty weapon''?}
\begin{itemize}
\item Married two-parent families \ : \ desirable outcomes for parents and children%$\genfrac{}{}{0pt}{1}{\Leftarrow}{\Rightarrow}$
\item Pro-marriage policies: child support, tax brackets, divorce regulations, etc
\item Can economic policies create successful marriages? Love vs money.
\end{itemize}
\end{frame}

@@

\frametitle{Kids-First Marriages are Less Stable}
\begin{enumerate}
\item{\textbf{Shotgun marriage}} --- couple marrying when woman gets pregnant
\item{\textbf{Kids-first marriage}} --- child \textit{before or in the year} of marriage
\end{enumerate}
\begin{itemize}
\item Around 1/4 marriages in the US today are kids-first
\item They have much higher divorce rates, especially among college graduates
\item Marginal in economic sense: could be sensitive to policies 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Roadmap}
\begin{enumerate}
\item Document the shotgun marriage and its outcomes empirically
\item Use these empirical relations to estimate a model capturing:
\begin{itemize}
\item dynamic bargaining and limited commitment in marriage
\item lifecycle, income and savings
\item fertility and child expenditures
\end{itemize}
\item Look at what drives decisions to marry and the differences in divorce
\begin{itemize}
\item what explains the education gradient?
\item can policies fix the difference or reduce the share?
\item what is the value of marriage in general?
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Summary of Results}
\begin{itemize}
\item 23\% of women had the first child before their first marriage (kids first)
\item They are 1.8 times more likely to be divorced than ``marriage first''
%\item College grads: 2.8 times more likely to be divorced
%\begin{itemize}
%\item though smaller kids first share (just 10\%)
%\end{itemize}
\item My structural model corrects selection and data imperfections
%\item Model results:
%\begin{itemize}
%\item Social stigma + threat of raising the child alone explains most of the difference
%\item Child support $\Rightarrow$ more single mothers, little effect on marriage decisions
%\item Being able to control fertility helps everyone
%\end{itemize}
\item Model answers: Love $>$ Money. In particular:
\begin{itemize}
\item Threat of raising the child alone explains most of the difference
\item Regulating marriages by policies is both hard and inefficient
\item Easier control over fertility fixes shotgun marriages and has bigger return
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Literature}
\begin{itemize}
\item Policy literature:
\begin{itemize}
\item Empirical: Alessina, Guiliano, 2005;  Tannenbaum, 2020; Rossin-Slater, 2017
\item Structural: Forester, 2020; Brown, Flinn, 2011
\item Focus on particular policies, ignore interactions, limited quantitative power
\end{itemize}
\item Lifecycle literature:
\begin{itemize}
\item Dynamic limited commitment for marriage and divorce: Mazzocco, 2007; Voena, 2015; Low et al., 2018; Shephard, 2019; Blasutto, 2020
\item Fertility during lifecycle: Sommer, 2015, Ejrnæs, Jørgensen, 2018
\item Link between divorce and fertility is missing
\end{itemize}
\item Empirics on shotgun marriages (demography and sociology):
\begin{itemize}
\item Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)
\item Mostly descriptive
\end{itemize}
\end{itemize}
%
%\begin{itemize}
%\item Empirical literature documents the scope of shotgun marriage around the world:
%\begin{itemize}
%\item Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)
%\item Mostly descriptive
%\end{itemize}
%\item Dynamic limited commitment models of divorce and marriage:
%\begin{itemize}
%\item Marcet, Marimon (2019), Mazzocco (2007) --- setting up the framework
%\item Voena (2015), Shephard (2019); Lafortune, Low  (2019) --- applied questions
%\item Mostly ignore or simplify interaction with fertility
%\end{itemize}
%\item Unplanned and teenage pregnancy literature:
%\begin{itemize}
%\item Ejrnæs, Jørgensen (2018) --- lifecycle model
%\item Card, Wise (1978) to Ashcraft et al (2013) --- lots of empirical papers
%\item Treat household structure as fixed or exogenous
%\end{itemize}
%\end{itemize}
\end{frame}

\section{Data}
\begin{frame}
\frametitle{Sources: ACS 2009--2017 and others}
\begin{itemize}
\item Main exercises and model estimation are done on American Community Survey
%\item Huge sample size and enough information to show many things 
\item Individual marriage and fertility history are observed partially
\begin{itemize}
\item Only the most recent marriage date --- restrict to ``married once'' when matters
\item Pick age $\in [21,40]$ to have good information on fertility
\end{itemize}
\item Smaller datasets with better histories for robustness and extra evidence:
\begin{itemize}
\item Survey of Income and Program Participation (SIPP)
\item More: NSFG and NSFH
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Quantification}
\framesubtitle{ACS 2009--2017, Women 21--40, married once}
\begin{center}
\begin{tabular}{l r r r r }
\hline
& \multicolumn{2}{c}{share of divorced if ... }&  \\
&  marriage-first & kids-first & (share of kids-first) &  \\\hline
\multicolumn{5}{l}{\textbf{All sample}} \\\hline
\textit{Cross-sectional share of divorced} &  9.9 & 18.1 & (26.1) \\
\textit{Divorced 5 years after marriage} &  5.1 & 14.3  & (24.6) \\
\textit{Divorced 10 years after marriage} & 10.4 & 22.8 & (24.4) \\\hline
\multicolumn{5}{l}{\textbf{Cross-sectional share of divorced for subsamples}} \\\hline
\textit{High school only} &  12.8 & 17.3 & (37.0) \\
\textit{Some college} & 14.3 & 21.0 & (31.3) \\
\textit{College or more} &   5.3 & 14.8 & (11.9) \\\hline
\textit{First birth before 25} & 15.5 & 19.9 & (41.8) \\
\textit{First birth at 25 or later} &  6.2 & 10.9 & (10.7)  \\\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Controls}
\begin{center}
%\vspace{-0.5cm}
\begin{tabular}{l r r r }
\multicolumn{4}{c}{\textit{Regression equation:} $\text{Divorced}_i = \Delta \cdot \text{kids-first}_i + \text{Controls}_i + \varepsilon_i$} \\\hline
\hline
& \multicolumn{2}{c}{Estimates}  \\
& Difference ($\Delta$) & (standard error)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, regression}} \\\hline
\textit{No controls (raw difference) }& 8.2 & (0.1)  \\
\textit{Demographic controls }& 5.7 & (0.1)  \\
\textit{Duration controls} &  5.8 & (0.2)  \\
\textit{Demographic + duration} &  4.8 & (0.2)  \\
\textit{Demographic + duration + income} & 4.1 & (0.3)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, propensity score matching}} \\\hline
\textit{Demographic controls} & 5.0 & (0.3)  \\
\textit{Demographic controls + income} & 4.9 & (0.2)  \\\hline
\multicolumn{4}{p{0.825\linewidth}}{\footnotesize \textit{Controls:} demographics: FE for age $\times$ education, race, state, survey year. Duration: FE for age $\times$ age of the first marriage and age $\times$ age of the first birth.  Income: 3rd degree polynomial of log gross income (if observed).}\\\hline\hline
\end{tabular}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Kids First Women Are Divorced More Often: Extra Evidence from SIPP}
\begin{center}
\begin{tabular}{l r r }
\hline
\multicolumn{3}{l}{\textbf{Including remarried people...}} \\\hline
& \multicolumn{2}{c}{share of divorced if ... }  \\
&  marriage-first & kids-first   \\\hline
\textit{Ever divorced} & 34.4 & 48.0  \\
\textit{Divorced if married once} &  12.4 & 19.3 \\
\textit{Remarried} &  22.0 & 28.7 \\\hline
\hline
\multicolumn{3}{l}{\textbf{Separating step and own children (2+ kids only)}} \\\hline
& share in sample & share of divorced   \\\hline
{Marriage-first} & 72.1 & 10.5 \\
{Kids-first-own} &  16.4 & 14.6 \\
{Kids-first-other} &  11.4 & 27.2 \\\hline
\multicolumn{3}{p{0.7\linewidth}}{\footnotesize \textit{Notes.} SIPP data, the cross-section of Wave 1, 2014, women 21+. Numbers are percents.}\\\hline\hline
\end{tabular}
\end{center}
\end{frame}

@@

\begin{frame}
\frametitle{Value Function: Female, Single}
\framesubtitle{$p^m_t$ --- probability of meeting, $p^p_t$ --- probability of unplanned pregnancy, $m_t$ --- agree to marry}
\begin{align*}
V^{f,s}_t(a,w) = \max\limits_{c,a'} \Bigg\{ \frac{c^{1-\sigma}}{1-\sigma} + \beta \cdot \E_t \Big\{  & 
%& \\ \text{ (did not meet anyone): \ } & 
(1-p^{\text{m}}_t)\cdot V_{t+1}^{f,s}(a',w') \\
\text{ \footnotesize (met and not pregnant): \ }& +  p^{\text{m}}_t\cdot (1- p^{p}_t) \cdot \left[ m^{np}_t \cdot V^{f,c}_{t+1}  + (1-m^{np}_t)\cdot  V^{f,s}_{t+1}\right] \\
\text{ \footnotesize (met and pregnant): \ } & + p^{\text{m}}_t \cdot p^{\text{p}}_t\cdot   \left[ m^p_t \cdot V^{f,ck}_{t+1}  + (1-m^p_t)\cdot  V^{f,\text{preg}}_{t+1}\right] \Big\} \Bigg\}, \\
\text{ \footnotesize (uncertain abortion access): } &  V^{f,\text{preg}}_{t+1} = p^{\text{access}} \cdot \max\{V^{f,sk}_{t+1},V^{f,s}_{t+1} - \phi_a\} \\
   & \hspace{1cm} + (1-p^{\text{access}})\cdot V^{f,sk}_{t+1},\\
\text{ \footnotesize (budget constraint): \ } & c + a' = R\cdot a  + w , \\
\text{ \footnotesize (wage evolution): \ } & \log w = z + \text{Trend}_t, \ \ z' = z + \varepsilon^{f}.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple, No Children}
\framesubtitle{$k_t$ --- give a birth, $d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{c}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,a',k} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay, no birth): \ } & + (1-d_t) \cdot (1-k_t) \cdot  \left[ \theta^f \cdot  V^{f,c}_{t+1} + \theta^m \cdot V^{m,c}_{t+1}  \right] \\
\text{ \footnotesize (stay, give a birth): \ } &  + (1-d_t) \cdot k_t \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right]  \\
\text{ \footnotesize (divorce): \ } & + d_t  \left[ \theta^f \cdot  \tilde{V}^{f,s}_{t+1} + \theta^m \cdot \tilde{V}^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint, rts): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + a' = R\cdot a + w_f + w_m}\\
\text{ \footnotesize (match quality evolution): \ } &{\normalsize \psi' = \psi + \varepsilon^{\psi}}\\
\text{ \footnotesize \textbf{(participation constraints)}: \ } &{\normalsize V^{f,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{f,s}_{t+1}, \ \ V^{m,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{m,s}_{t+1}.}\\
\end{align*}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple and Child}
\framesubtitle{$d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{ck}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,x,l_f,a'} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \\ \text{ \footnotesize (child is a public good): \ }  & \phi + \alpha \cdot \frac{Q^{1-\sigma}}{1-\sigma} +  \beta \cdot \E_t \Big\{   \\
\text{ \footnotesize (stay): \ } & + (1-d_t) \cdot  \left[ \theta^f \cdot  V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}  \right] \\
\text{ \footnotesize (divorce): \ } & + d_t \cdot \left[ \theta^f \cdot  \tilde V^{f,sk}_{t+1} + \theta^m \cdot \tilde V^{m,s}_{t+1}  \right]  \Big\} \Bigg\}\\
\text{ \footnotesize (budget constraint): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + x + a' = R\cdot a + w_f\cdot l_f + w_m}\\
\text{ \footnotesize (child consumption flow): \ } &{\normalsize Q = \left[x^\lambda + \kappa \cdot (1-l_f)^\lambda\right]^{\frac1\lambda}, \ \ \ Q\geq \bar{Q}} \text{ \footnotesize  (subsistence constraint) } \\
\text{ \footnotesize (female skills $\downarrow$ if not working): \ } &{\normalsize \log w_f = z_f + \text{Trend}_{f,t}, \ \ z'_f = z_f + \varepsilon^f - \delta(l^f)}\\
\end{align*}
\end{frame}

@@

>>>>

%\setbeamertemplate{footline}[frame number]

\setbeamertemplate{footline}{% 
  \hfill% 
  \usebeamercolor[fg]{page number in head/foot}% 
  \usebeamerfont{page number in head/foot}% 
  \insertframenumber%
  %\,/\,\inserttotalframenumber
  \kern1em\vskip2pt% 
}

@@

\begin{frame}
\frametitle{Can Economic Policies Create Successful Marriages?}
\begin{itemize}
\item Reproduction of poverty is a motivation
\item Married two-parent families \ $\genfrac{}{}{0pt}{1}{\Leftarrow \text{ or}}{\Rightarrow \text{  ?}}$ \ desirable outcomes for parents and children%
\item Pro-marriage policies: child support, tax brackets, divorce regulations, etc
\item I propose a measurable dimension of marriage heterogeneity
\end{itemize}
\end{frame}

@@

\frametitle{Kids-First Marriages are Closer to The Threshold}
\begin{enumerate}
\item{\textbf{Kids-first couples}} --- child \textit{before or in the year} of marriage
\item{\textbf{Marriage-first couples}} --- child \textit{at the following year or later}
\end{enumerate}
\begin{itemize}
\item Around 1/4 of childbearing marriages are kids-first
\item They have much higher divorce rates, especially among college graduates
\item Natural distinction to study marriage quality
%\item Marginal in economic sense: sensitive to policies 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Roadmap}
\begin{enumerate}
\item Document the kids-first marriage and its outcomes empirically
\item Use these empirical relations to estimate a model capturing:
\begin{itemize}
\item dynamic bargaining and limited commitment in marriage
\item lifecycle, income and savings
\item fertility and child expenditures
\end{itemize}
\item Use it to correct the data imperfections and answer the questions:
\begin{itemize}
\item Why do the kids-first couples divorce more?
\item Can policies fix the difference or reduce the share?
\item How to promote \emph{good} marriages?
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Conceptual Framework}
\begin{itemize}
\item Marriages differ in match quality $\psi$ (amount of love)
\item Potential spouses observe their $\psi$ and marry if $\psi > \psi_m$ (\textit{agreement threshold})
\item Match quality evolves, and couple divorces if later $\psi < \psi_d$ (\textit{divorce threshold})
\item Lower $\psi$ implies higher risk of divorce
\item Structural model recovers \textit{distribution} of $\psi$ from the risks of divorce
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Summary of Results}
\begin{enumerate}
\item Differences in divorce:
\begin{itemize}
\item Kids before marriage lower the agreement threshold, divorce chances increase by 1/3
\item 40--50\% of the kids-first marriages happen \emph{because} of kids
\item Risks of being a single mother and social stigma are the main factors
\end{itemize}
\item Implications for policies:
\begin{itemize}
\item Pushing people to marry more is welfare-decreasing for everyone
\item Making divorce harder has some winners, but people lose on average
\item Child support has limited effects on marriage formation, but increases fertility, stimulates divorce and may hurt children
\end{itemize}
%\item 23\% of women had the first child before their first marriage (kids first)
%\item They are 1.8 times more likely to be divorced than ``marriage first''
%\item College grads: 2.8 times more likely to be divorced
%\begin{itemize}
%\item though smaller kids first share (just 10\%)
%\end{itemize}
%\item My structural model corrects selection and data imperfections
%\item Model results:
%\begin{itemize}
%\item Social stigma + threat of raising the child alone explains most of the difference
%\item Child support $\Rightarrow$ more single mothers, little effect on marriage decisions
%\item Being able to control fertility helps everyone
%\end{itemize}
%\item Model answers: Love $>$ Money. In particular:
%\begin{itemize}
%\item Threat of raising the child alone explains most of the difference
%\item Regulating marriages by policies is both hard and inefficient
%\item Easier control over fertility fixes shotgun marriages and has bigger return
%\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Literature}
\begin{itemize}
\item Empirics on shotgun marriages (demography and sociology):
\begin{itemize}
\item {\small  \color{gray}Gibson-Davis et al (2016), Guzzo (2009),  Nuevo-Chiquero (2014)}
\item I can recover the causal content of these numbers
\end{itemize}
\item Lifecycle models:
\begin{itemize}
\item Dynamic limited commitment for marriage and divorce: {\small  \color{gray}Mazzocco, 2007; Voena, 2015; Low et al., 2018; Shephard, 2019; Blasutto, 2020}
\item Fertility during lifecycle: {\small  \color{gray}Sommer, 2015, Ejrnæs, Jørgensen, 2018}
\item I am the first to estimate a marriage/divorce model with fertility choice
\end{itemize}
\item Policy literature:
\begin{itemize}
\item Empirical: {\small \color{gray}Alessina, Guiliano, 2005;  Tannenbaum, 2020; Rossin-Slater, 2017}
\item Structural: {\small  \color{gray}Forester, 2020; Brown, Flinn, 2011; Kennes, Knowles, 2020}
\item I am the first to consider creation and dissolution of shotgun marriages structurally
\end{itemize}
\end{itemize}
\end{frame}

\section{Empirical Motivation}
\begin{frame}
\frametitle{Sources: ACS 2009--2017 and others}
\begin{itemize}
\item Main exercises and model estimation are done on American Community Survey
%\item Huge sample size and enough information to show many things 
\item Individual marriage and fertility history are observed partially
\begin{itemize}
\item Only the most recent marriage date --- restrict to ``married once'' when matters
\item Pick age $\in [21,40]$ to have good information on fertility
\end{itemize}
\item Smaller datasets with better histories for robustness and extra evidence:
\begin{itemize}
\item Survey of Income and Program Participation (SIPP)
\item More: NSFG and NSFH
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=dt_more]
\frametitle{Kids First Women Are Divorced More Often: More Evidence}
\begin{itemize}
\item The divorce pattern is present in {all subgroups}
\item The pattern is \textit{sharper} for {college graduates} or those with later births
\item Controlling for many confounders explains no more than a half
\item Step-children drive relatively small part of the result
\end{itemize}
\vspace{1cm}
\begin{tabular}{c c c c}
More results: & \hyperlink{emp-table}{\beamerbutton{subgroups}} & \hyperlink{emp-controls}{\beamerbutton{controls}} & \hyperlink{emp-step}{\beamerbutton{step-children}} 
\end{tabular}

@@

\begin{frame}
\frametitle{Value Function: Female, Single}
\framesubtitle{$p^m_t$ --- probability of meeting, $p^p_t$ --- probability of unplanned pregnancy, $m_t$ --- agree to marry}
\begin{align*}
V^{f,s}_t(a,w) = & \max\limits_{c,a'} \Bigg\{ \frac{c^{1-\sigma}}{1-\sigma} + \beta \cdot \E_t  \text{\textbf{Transitions}}_{t+1} \Bigg\}  \\
\text{ \footnotesize (budget constraint): \ } & c + a' = R\cdot a  + w , \\
\text{ \footnotesize (wage evolution): \ } & \log w = z + \text{Trend}_t, \ \ z' = z + \varepsilon^{f}.
\end{align*}
\textbf{Transitions}:
\begin{itemize}
\item \footnotesize No partner met: stay single, $V^{f,s}_{t+1}$, random with probability $1-p^{\text{meet}}_t$
\item \footnotesize Met, no unplanned pregnancy, random with probability $p^{\text{meet}}_t\cdot (1-p^{\text{preg}}_t)$: 
\begin{itemize}
\item \footnotesize Agree to marry: enter couple without a child, $V^{f,c}_{t+1}$
\item\footnotesize  Disagree, stay single: $V^{f,s}_{t+1}$
\end{itemize}
\item \footnotesize Met, no unplanned pregnancy, random with probability $p^{\text{meet}}_t\cdot p^{\text{preg}}_t$: 
\begin{itemize}
\item \footnotesize Agree to marry: enter couple with a child, $V^{f,ck}_{t+1}$
\item \footnotesize Disagree, \textit{pick} $\max\{V^{f,s}_{t+1} - \phi_a, V^{f,sk}_{t+1}\}$ \textit{if abortion is possible ($p^a$), otherwise} $V^{f,sk}_{t+1}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple, No Children}
\framesubtitle{$k_t$ --- give a birth, $d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\begin{align*}
V^{c}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,a',k} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \beta \cdot \E_t \text{\textbf{Transitions}}_{t+1} \Bigg\}   \\
\text{ \footnotesize (budget constraint, rts): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + a' = R\cdot a + w_f + w_m}\\
\text{ \footnotesize (match quality evolution): \ } &{\normalsize \psi' = \psi + \varepsilon^{\psi}}\\
\text{ \footnotesize \textbf{(participation constraints)}: \ } &{\normalsize V^{f,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{f,s}_{t+1}, \ \ V^{m,c\bullet}_{t+1} (\cdots,\theta') \geq \tilde{V}^{m,s}_{t+1}.}
\end{align*}
\textbf{Transitions}:
\begin{itemize}
\item  \footnotesize Stay together, no birth, choice: $\theta^f \cdot V^{f,c}_{t+1} + \theta^m \cdot V^{m,c}_{t+1}$
\item  \footnotesize Stay together, try to give a birth, choice:
\begin{itemize}
\item  \footnotesize  \textit{If successful} ($p^{\text{birth}}_t$):  $\theta^f\cdot V^{f,ck}_{t+1} + \theta^m\cdot V^{m,ck}_{t+1}$, \textit{otherwise} $\theta^f\cdot V^{f,c}_{t+1} + \theta^m\cdot V^{m,c}_{t+1}$
\end{itemize}
\item  \footnotesize  Divorce, choice: $\theta^f \cdot \tilde V^{f,s}_{t+1} + \theta^m \cdot \tilde V^{m,s}_{t+1}$ (property division + divorce costs inside)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Value Function: Couple and Child}
\framesubtitle{$d_t$ --- get a divorce, $\tilde{V}$ --- values at divorce}
\vspace{-1cm}
\begin{align*}
V^{ck}_t(a,w_f,w_m,\psi,\theta) = \max\limits_{c_f,c_m,x,l_f,a'} \Bigg\{ & \theta^f \cdot \frac{c_f^{1-\sigma}}{{1-\sigma}}  + \theta^m \cdot \frac{c_m^{1-\sigma}}{1-\sigma}  + \psi + \\ \text{ \footnotesize (child is a public good): \ }  & \phi + \alpha \cdot \frac{Q^{1-\sigma}}{1-\sigma} +  \beta \cdot \E_t \text{\textbf{Transitions}}_{t+1}  \Bigg\} \\
\text{ \footnotesize (budget constraint): \ } &{\normalsize \left[c_f^{1+\rho} + c_m^{1+\rho}\right]^{\frac1{1+\rho}} + x + a' = R\cdot a + w_f\cdot l_f + w_m}\\
\text{ \footnotesize (child consumption flow): \ } &{\normalsize Q = \left[x^\lambda + \kappa \cdot (1-l_f)^\lambda\right]^{\frac1\lambda}}, \\
\text{ \footnotesize  (subsistence constraint): } & Q\geq \bar{Q} \\
\text{ \footnotesize (female skills $\downarrow$ if not working): \ } &{\normalsize \log w_f = z_f + \text{Trend}_{f,t}, \ \ z'_f = z_f + \varepsilon^f - \delta(l^f)}
\end{align*}
\textbf{Transitions}:
\begin{itemize}
\item  \footnotesize Stay together, choice: $\theta^f \cdot V^{f,ck}_{t+1} + \theta^m \cdot V^{m,ck}_{t+1}$
\item  \footnotesize Divorce, choice: $\theta^f \cdot \tilde V^{f,sk}_{t+1} + \theta^m \cdot \tilde V^{m,s}_{t+1}$ (property division + divorce costs + child support inside)
\end{itemize}
\end{frame}

@@

\begin{frame}[plain,label=emp-table]
\frametitle{Kids First Women Are Divorced More Often: Quantification}
\framesubtitle{ACS 2009--2017, Women 21--40, married once}
\begin{center}
\begin{tabular}{l r r r r }
\hline
& \multicolumn{2}{c}{share of divorced if ... }&  \\
&  marriage-first & kids-first & (share of kids-first) &  \\\hline
\multicolumn{5}{l}{\textbf{All sample}} \\\hline
\textit{Cross-sectional share of divorced} &  9.9 & 18.1 & (26.1) \\
\textit{Divorced 5 years after marriage} &  5.1 & 14.3  & (24.6) \\
\textit{Divorced 10 years after marriage} & 10.4 & 22.8 & (24.4) \\\hline
\multicolumn{5}{l}{\textbf{Cross-sectional share of divorced for subsamples}} \\\hline
\textit{High school only} &  12.8 & 17.3 & (37.0) \\
\textit{Some college} & 14.3 & 21.0 & (31.3) \\
\textit{College or more} &   5.3 & 14.8 & (11.9) \\\hline
\textit{First birth before 25} & 15.5 & 19.9 & (41.8) \\
\textit{First birth at 25 or later} &  6.2 & 10.9 & (10.7)  \\\hline
\end{tabular}
\end{center}
\hyperlink{dt_more}{\beamerbutton{back}} 
\end{frame}

\begin{frame}[plain,label=emp-controls]
\frametitle{Kids First Women Are Divorced More Often: Controls}
\begin{center}
%\vspace{-0.5cm}
\begin{tabular}{l r r r }
\multicolumn{4}{c}{\textit{Regression equation:} $\text{Divorced}_i = \Delta \cdot \text{kids-first}_i + \text{Controls}_i + \varepsilon_i$} \\\hline
\hline
& \multicolumn{2}{c}{Estimates}  \\
& Difference ($\Delta$) & (standard error)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, regression}} \\\hline
\textit{No controls (raw difference) }& 8.2 & (0.1)  \\
\textit{Demographic controls }& 5.7 & (0.1)  \\
\textit{Duration controls} &  5.8 & (0.2)  \\
\textit{Demographic + duration} &  4.8 & (0.2)  \\
\textit{Demographic + duration + income} & 4.1 & (0.3)  \\\hline
\multicolumn{4}{l}{\textbf{All sample, propensity score matching}} \\\hline
\textit{Demographic controls} & 5.0 & (0.3)  \\
\textit{Demographic controls + income} & 4.9 & (0.2)  \\\hline
\multicolumn{4}{p{0.825\linewidth}}{\footnotesize \textit{Controls:} demographics: FE for age $\times$ education, race, state, survey year. Duration: FE for age $\times$ age of the first marriage and age $\times$ age of the first birth.  Income: 3rd degree polynomial of log gross income (if observed).}\\\hline\hline
\end{tabular}
\end{center}
\vspace{-0.5cm}
\hyperlink{dt_more}{\beamerbutton{back}} 
\end{frame}

\begin{frame}[plain,label=emp-step]
\frametitle{Kids First Women Are Divorced More Often: Extra Evidence from SIPP}
\begin{center}
\begin{tabular}{l r r }
\hline
\multicolumn{3}{l}{\textbf{Including remarried people...}} \\\hline
& \multicolumn{2}{c}{share of divorced if ... }  \\
&  marriage-first & kids-first   \\\hline
\textit{Ever divorced} & 34.4 & 48.0  \\
\textit{Divorced if married once} &  12.4 & 19.3 \\
\textit{Remarried} &  22.0 & 28.7 \\\hline
\hline
\multicolumn{3}{l}{\textbf{Separating step and own children (2+ kids only)}} \\\hline
& share in sample & share of divorced   \\\hline
{Marriage-first} & 72.1 & 10.5 \\
{Kids-first-own} &  16.4 & 14.6 \\
{Kids-first-other} &  11.4 & 27.2 \\\hline
\multicolumn{3}{p{0.7\linewidth}}{\footnotesize \textit{Notes.} SIPP data, the cross-section of Wave 1, 2014, women 21+. Numbers are percents.}\\\hline\hline
\end{tabular}
\end{center}
\hyperlink{dt_more}{\beamerbutton{back}} 
\end{frame}
